### 2016-09-22 paper.tex
changes

\newtheorem{cor}{Corallary}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

@@

\newcommand{\zero}{\text{\textbf{0}}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\W}{\mathcal{\hat W}}
\newcommand{\Waff}{\mathcal{\hat A}}
\newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
\newcommand{\Wtave}{{\mathcal{W}^{ave,*}}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\theta}
\newcommand{\wkl}{\hat\w^{kl}}
\newcommand{\wreopt}{\hat\w^{reopt}}
\newcommand{\wave}{\hat\w^{ave}}
\newcommand{\wtave}{\E\hat\w^{ave}}
\newcommand{\waver}{\hat\w^{ave,r}}
\newcommand{\wboot}{\hat\w^{boot}}
\newcommand{\wmle}{\hat\w^{mle}}
\newcommand{\wmler}{\hat\w^{mle,r}}
\newcommand{\wstar}{{\w^{*}}}
\newcommand{\wq}{\hat\w^{q}}
\newcommand{\wqstar}{\hat\w^{q^*}}
\newcommand{\dist}{\mathcal{D}}

@@

@article{spokoiny2012parametric,
  title={Parametric estimation. Finite sample theory},
  author={Spokoiny, Vladimir},
  journal={The Annals of Statistics},
  volume={40},
  number={6},
  pages={2877--2909},
  year={2012},
  publisher={Institute of Mathematical Statistics}
}

@@

\begin{abstract}
\end{abstract}

@@

Let $\Y\subseteq\mathbb{R}$ be the space of response variables,
$X\subseteq\mathbb{R}^d$ be the space of the covariates,
and $\Theta\subseteq\mathbb{R}^d$ be the parameter space.
We assume a linear model.
The log-likelihood of data point $(\x,y)\in\X\times\Y$ given the model's true parameter $\wstar\in\Theta$ is denoted by $f(y,\trans\x\wstar)$.
Our analysis in Section \ref{sec:anal} places very mild restrictions on $f$.
In particular, $f$ need not be convex.
%For our experiments in Section \ref{sec:exp}, however, we restrict ourselves to the likelihood for logistic regression
%\begin{equation}
%f(y,\trans\x\w) =
%\end{equation}
Let $Z\subset\X\times\Y$ be a dataset of $mn$ i.i.d. observations.
Then the maximum likelihood estimator is
\begin{equation}
%\wmle=\argmax_\w \sum_{i=1}^{nm} g(y_i-\trans\x_i\w)
\wmle=\argmax_\w \sum_{(\x,y)\in Z} f(y,\trans\x\w)
\end{equation}
Assume that $Z$ has been partitioned onto $m$ machines so that each machine $i$ has dataset $Z_i$ of size $n$, and all the $Z_i$ are disjoint.
Then each machine calculates the local estimator
\begin{equation}
\wmle_i = \argmax_\w \sum_{(\x,y) \in Z_i} f(y,\trans\x\w)
\end{equation}
Solving for $\wmle_i$ requires no communication with other machines.
Our goal is to provide a function $q : \Theta^m \to \Theta$ that ``merges'' these estimates.
The resulting merged estimate should have error as close to $\wmle$ as possible.

@@

%The averaging estimator is defined as
%\begin{equation}
%\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
%\end{equation}
%Then our estimator is defined as
%\begin{equation}
%\begin{aligned}
%\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x (\hat W\alpha + \wave)\right)
%\\
%\wreopt &= \hat W \hat\alpha + \wave
%\end{aligned}
%\end{equation}
%where $Z^{reopt}$ is a small dataset of size $O(m)$.
Define the matrix $\hat W : \mathbb{R}^{d\times m}$ to have $i$th column equal to $\wmle_i$.
Let $n^{reopt}$ be an integer less than $n$,
and $Z^{reopt}_i$ be a subset of $Z_i$ randomly sampled without replacement.
Then our estimator is defined as
\begin{equation}
\begin{aligned}
%Z_i^{proj} &= \{(y,\x\hat W) : (y,\x) \in Z_i^{reopt}\}
%\\
\hat\alpha &= \argmax_\alpha \sum _{(\x,y)\in Z^{reopt}} f\left(y,\trans\x \hat W\alpha \right)
\\
\wreopt &= \hat W \hat\alpha
\end{aligned}
\end{equation}
%In the analysis of this estimator, we assume that $Z^{reopt}$ is independent of the $\wmle_i$;
%that is, the data points in $Z^{reopt}$ were not used in the single machine optimizations.
%In practice, however, this restriction is not necessary.

%Notice that this estimator requires two rounds of communication.
There are two ways to implement this estimator.
The simplest is with two rounds of communication.
In the first round, each machine broadcasts its value of $\wmle_i$ to every other machine.
In the second round, each machine sends $Z_i^{proj}$ to a predesignated machine.
This predesignated machine performs the second optimization and outputs the final answer.
The full procedure is shown in Algorithm \ref{alg:distributed} above.
There is an alternative implementation if there is a common set of data points $Z^{common}$ possessed by each machine in addition to $Z_i$.

Our estimator has the following intuitive interpretation.
Let $\W$ be the column space of $\hat W$.
We use the $\wmle_i$ estimates from the first round of communication to create a subspace $\Wave=\affspan\{\wave\}\cup{\{\wmle\}}_{i=1}^m$.
Then in the second round of optimization, we project all of our data points down into this subspace.
Because this subspace is small, we need only a small number of data points to get a good estimate of the optimum.
Because of the way we constructed $\Wave$,
the optimum in $\Wave$ will be close to the true optimum.
%The true parameter $\wstar$ should be close to this subspace.
%In the first round of optimization (determining the $\wmle_i$s),
%we're actually finding a subspace
%The estimator $\wreopt$ lives in a subs

\begin{figure}
\hspace{-0.1in}
\begin{tikzpicture}
    [
    dot/.style = {minimum width=0.15cm,inner sep=0pt,line width=0pt,fill,circle,black,font=\small}
    ]
\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.5,-0.5) (-1.05,1) (-0.1,-0.1) (0.75,-0.5) (0.45,-0.7) };
\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.85,-0.65) (-1.3,1.55) (0.25,0) (2.5,-0.5) (1,-0.95)};
\draw[gray] plot [smooth cycle,tension=1] coordinates {(-1.15,-1) (-1.5,2) (1,0) (5,0) (3,-0.95)};

\draw[thick] (-2.2,2.15) -- (5.5,1);

\node[dot] (wstar) at (-0.25,-0.25) {};
\node at (0,-0.45) {$\wstar$};

@@

%\node[dot] (wavestar) at (4,0.5) {};
%\node                 at (4.5,0.3) {$\E\wave$};
\node[dot] (wave) at (3.8,1.25) {};
\node at (4.0,1.6) {$\wave$};
\node[dot] (wave) at (2.9,1.40) {};
\node at (3.1,1.75) {$\wmle_1$};
\node[dot] (wave) at (4.7,1.10) {};
\node at (4.9,1.45) {$\wmle_2$};
%\node[dot] (wave) at (3.8,1.25) {};

\node[dot] at (-1.35,2.03) {};
\node at (-1.1,2.4) {$\wreopt$};

\node at (4,-1.25) {$f(y,\trans\x\w)$};
\node at (5.75,1) {$\W$};

\end{tikzpicture}
\caption{
    Our method performs a second round of optimization to find the best parameter vector in $\W$.
    Since $\wave\in\W$, our estimator $\wreopt$ is guaranteed to have higher true likelihood with high probability.
}
\label{fig:contour}
\end{figure}

\begin{algorithm}[t]
\caption{Estimation in two rounds}
\label{alg:distributed}
\begin{algorithmic}
\State Each machine $i$ independently
\State ~~~~~loads dataset $Z_i$
\State ~~~~~calculates $\wmle_i$
\State All machines broadcast $\wmle_i$ to all other machines
\State Each machine $i$ independently
\State ~~~~~constructs $\hat W$
\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
\State ~~~~~calculates $\trans\x\hat W$
\State All machines broadcast $Z^{reopt}_i$ to a master machine
\State The master calculates $\wreopt$
\end{algorithmic}
%\label{fig:alg}
\end{algorithm}

\begin{algorithm}[t]
\caption{Estimation with common data}
\label{alg:distributed}
\begin{algorithmic}
\State Each machine $i$ independently
\State ~~~~~loads dataset $Z_i$
\State ~~~~~calculates $\wmle_i$
\State All machines broadcast $\wmle_i$ to all other machines
\State Each machine $i$ independently
\State ~~~~~constructs $\hat W$
\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
\State ~~~~~calculates $\trans\x\hat W$
\State All machines broadcast $Z^{reopt}_i$ to a master machine
\State The master calculates $\wreopt$
\end{algorithmic}
%\label{fig:alg}
\end{algorithm}

@@

\subsection{Notation}

@@

\subsection{In the Normal Regime}
%\label{sec:normal}
%
%We say that a random vector $X$ is \emph{sub-Gaussian} with mean $\mu$ and covariance $\Sigma$
%if there exists a constant $c>0$ such that
%\begin{equation}
%\prob{\ltwo{\Sigma(X-\mu)} > t} \le c \exp(-t^2)
%\end{equation}
%We denote this relationship in symbols as $X \lesssim \normal{\mu}{\Sigma}$.
%
%\begin{theorem}
%We have that with probability at least
%\begin{equation}
%\wmle-\wstar \lesssim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
%\end{equation}
%In the limit as $n\to\infty$, we have the stronger statement that
%\begin{equation}
%\wmle-\wstar \sim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
%\end{equation}
%\end{theorem}

@@

We assume that for every estimator $\hat\w$ trained on $n$ data points,
\begin{equation}
\label{eq:assumption}
\hat\w - \E\hat\w \sim \frac{1}{\sqrt n}\normal{\zero}{\I^{-1}_{\E\hat\w}}
%\hat\w \sim \normal{\E\hat\w}{\frac{1}{n}I^{-1}_{\E\hat\w}}
\end{equation}
%
%Throughout Section \ref{sec:normal} we assume that $\wave_i$ is normally distributed with mean $\wtave$ and covariance $\frac{1}{n}I^{-1}_{\wtave}$.
%\begin{equation}
%\label{eq:assumption}
%\wmle_i \sim \normal{\E\wmle_i}{\frac{1}{n}I^{-1}_{\E\wmle_i}}
%\end{equation}
%Here, $I_{\wtave}$ is the Fisher information matrix at point $\wtave$.
Here, $\I_{\E\hat\w}$ is the Fisher information matrix for parameter $\E\hat\w$.
This assumption is motivated by the central limit theorem for maximum likelihood estimation.
%There are many versions of this theorem.
Theorem 7.5.2 of \cite{lehmann1999elements} is particularly simple version and reproduced below.
\begin{theorem}
\label{thm:clt}
Assume the following mild regularity conditions:
\begin{enumerate}[noitemsep,topsep=0pt]%,label=(\itshape\roman*)]
\item all data points in $Z_i$ are i.i.d.;
\item the parameter space $\Theta$ is open;
\item the model is identifiable and consistent;
\item the matrix $\I_{\wtave}$ is non-singular;
\item the support set $A = \{(y,\x) : f(y,\x;\theta) > 0\}$ is independent of $\theta$;
and \item for all $(y,\x)\in A$, $f(y,\x;\theta)$ is three times differentiable with respect to $\theta$,
and the third derivative is continuous.
\end{enumerate}
Then in the limit as $n\to\infty$,
\begin{equation*}
\wave_i \law \wtave + \frac{1}{\sqrt n} \normal{0}{\I^{-1}_\wstar}
\end{equation*}
Here the notation $\law$ denotes convergence in law,
which is the standard notion of convergence for a sequence of probability distributions.
\end{theorem}

@@

There are many generalizations of Theorem \ref{thm:clt}.
For example, Theorem 4.2 in \cite{spokoiny2012parametric} provides finite sample bounds with explicit dependence on the degree of model misspecification, optimization error, and the lack of independence in the data variables.
Because Spokoiny's theorem is non-asymptotic, the theorem that the estimator is a sub-Gaussian random variable.

@@

Note that this reduction in variance is essentially optimal.
By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
we can show that the variance of the MLE trained on all of the data is
\begin{equation}
\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
\end{equation}
The only difference is the location of the Fisher information matrix.

@@

\begin{proof}
This proof follows the same form as the proof for Theorem \ref{thm:wave}.
We first decompose the error into a bias and variance component;
then we bound each component separately.
To decompose the error, we need to introduce the following affine space.
\begin{align}
\Waff&=\affspan \{\wmle_i\}_{i=1}^m
\\
\Waff_{\E}&=\affspan \{\wmle_i - \wave + \E\wave\}_{i=1}^m
\end{align}
We now have that
\begin{align}
\ltwo{\wstar-\proj\W\wstar}
&\le
\ltwo{\wstar-\proj\Waff\wstar}
\\
&\le
\ltwo{\wstar-\proj{\Waff_\E}\wstar}
%\\&
+
\ltwo{\proj{\Waff_\E}\wstar-\proj\Waff\wstar}
%\\&+
%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
\end{align}
The first line above follows because $\Waff\subseteq\W$,
and the second by the triangle inequality.

%Notice that $\Waff\subseteq\W$, so $\ltwo{\wstar-\proj\W\wstar} \le \ltwo{\wstar-\proj\Waff\wstar}$.
%\begin{equation}
%\Wtave=\{\x - \wave + \wtave : \x\in\Wave\}
%\end{equation}
%We have by the triangle inequality that
%\begin{equation}
%\begin{split}
%\ltwo{\wstar-\proj\Wave\wstar}
%&\le
%\ltwo{\wstar-\proj\Wtave\wstar}
%\\&+
%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
%\end{split}
%\end{equation}
The left term above can be thought of as akin to the bias of the estimator $\proj\Wtave\wstar$,
and the right term can be thought of as akin to the variance.

>>>>

\newtheorem{cor}{Corallary}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{assumption}{Condition}

@@

\newcommand{\zero}{\text{\textbf{0}}}
\newcommand{\Zproj}{Z^{\textit{proj}}}
\newcommand{\Zreopt}{Z^{\textit{owa}}}
\newcommand{\nreopt}{n^{\textit{owa}}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\W}{{\hat \Theta^{\textit{owa}}}}
\newcommand{\Waff}{\mathcal{\hat A}}
\newcommand{\WaffE}{{\mathcal{\hat A}_\E}}
\newcommand{\Wave}{{\mathcal{\hat W}^{ave}}}
\newcommand{\Wtave}{{\mathcal{W}^{ave,*}}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\theta}
\newcommand{\wkl}{\hat\w^{kl}}
\newcommand{\ahat}{\hat\alpha}
\newcommand{\afull}{\ahat^{\textit{full}}}
\newcommand{\wreopt}{\hat\w^{owa}}
\newcommand{\wreoptfull}{\hat\w^{\textit{owa,full}}}
\newcommand{\wave}{\hat\w^{ave}}
\newcommand{\wtave}{\E\hat\w^{ave}}
\newcommand{\waver}{\hat\w^{ave,r}}
\newcommand{\wboot}{\hat\w^{boot}}
\newcommand{\wmle}{\hat\w^{mle}}
\newcommand{\wmler}{\hat\w^{mle,r}}
\newcommand{\wstar}{{\w^{*}}}
\newcommand{\wq}{\hat\w^{q}}
\newcommand{\wqstar}{\hat\w^{q^*}}
\newcommand{\dist}{\mathcal{D}}

@@

@article{spokoiny2012parametricestimation,
  title={Parametric estimation. Finite sample theory},
  author={Spokoiny, Vladimir},
  journal={The Annals of Statistics},
  volume={40},
  number={6},
  pages={2877--2909},
  year={2012},
  publisher={Institute of Mathematical Statistics}
}

@@

\begin{abstract}
We present a distributed learning algorithm with low communication complexity.
Our algorithm is an easy to implement modification of the standard averaging method.
In the averaging method, each machine in a cluster first solves the learning problem on a subset of the data;
the resulting parameter vectors are then averaged together.
In our method, we perform a second round of optimization to find the \emph{optimal weighted average} of the parameter vectors.
This second optimization happens in a low dimensional space,
and so it adds negligible computational and communication costs.
%Given $m$ machines (each with $n$ data points of dimension $d$),
%we show that our algorithm reduces bias at a rate of $O(\frac{1-m/d}{n})$.
Whereas the averaging method reduces only the variance of the estimator,
our method reduces both the bias and the variance.
Our analysis is more general than the analysis of similar communication efficient algorithms that also reduce the bias.
Notably, we do not require the likelihood function to be convex.
In practice,
this bias reduction makes our algorithm significantly less sensitive to the amount of regularization.
This has the practical benefit of greatly speeding up the model selection process.
\end{abstract}

@@

Let $\Y\subseteq\mathbb{R}$ be the space of response variables,
$X\subseteq\mathbb{R}^d$ be the space of the covariates,
and $\Theta\subseteq\mathbb{R}^d$ be the parameter space.
We assume a linear model.
The log-likelihood of data point $(\x,y)\in\X\times\Y$ given the model's true parameter $\wstar\in\Theta$ is denoted by $f(y,\trans\x\wstar)$.
Our analysis in Section \ref{sec:anal} places very mild restrictions on $f$.
In particular, $f$ need not be convex.
%For our experiments in Section \ref{sec:exp}, however, we restrict ourselves to the likelihood for logistic regression
%\begin{equation}
%f(y,\trans\x\w) =
%\end{equation}
Let $Z\subset\X\times\Y$ be a dataset of $mn$ i.i.d. observations.
Then the maximum likelihood estimator is
\begin{equation}
%\wmle=\argmax_\w \sum_{i=1}^{nm} g(y_i-\trans\x_i\w)
\wmle=\argmax_\w \sum_{(\x,y)\in Z} f(y,\trans\x\w)
\end{equation}
Assume that $Z$ has been partitioned onto $m$ machines so that each machine $i$ has dataset $Z_i$ of size $n$, and all the $Z_i$ are disjoint.
Then each machine calculates the local estimator
\begin{equation}
\wmle_i = \argmax_\w \sum_{(\x,y) \in Z_i} f(y,\trans\x\w)
\end{equation}
Solving for $\wmle_i$ requires no communication with other machines.
%Our goal is to provide a function $q : \Theta^m \to \Theta$ that ``merges'' these estimates.
%The resulting merged estimate should have error as close to $\wmle$ as possible.
A non-interactive distributed estimator merges these $\wmle_i$ into a single improved estimate.
The simplest example is the averaging estimator
\begin{equation}
\wave = \frac{1}{m}\sum_{i=1}^m \wmle_i
\end{equation}
This estimator is well studied, and in Section \ref{sec:anal} we comprae this previous work to our own.
Here we briefly note that $\wave$ is known to reduce the variance of an estimator but not its bias.
Our goal is to design an estimator that reduces both variance and bias.

@@

In this section we present a modification to the averaging estimator called the \emph{Optimal Weighted Average} (OWA).
OWA uses a second round of optimization to calculate the optimal linear combination of the $\wmle_i$s.
%Using these optimal weights, we reduce bias at a rate of $O(\sqrt{1-m/d})$.
This second optimization occurs over a small fraction of the dataset,
so its computational and communication cost is negligible.

To motivate our estimator, we first describe a simpler estimator that uses the entire dataset for the second round of optimization.
%The idea of our proposed solution is to find the best linear combination of the $\wmle_i$s using a second round of optimization.
Define the matrix $\hat W : \mathbb{R}^{d\times m}$ to have $i$th column equal to $\wmle_i$.
Now consider the estimator
\begin{equation}
\wreoptfull = \hat W \afull
\end{equation}
where
\begin{equation}
\afull = \argmax_\alpha \sum _{(\x,y)\in Z} f\left(y,\trans\x \hat W \alpha \right)
\end{equation}
%This is the maximum likelihood estimator when we restrict our
Notice that $\wreoptfull$ is the maximum likelihood estimator when the parameter space $\Theta$ is restricted $\W = \vecspan \{\wmle_i\}_{i=1}^m$.
Figure \ref{fig:contour} shows graphically that no other estimator in $\W$ can have higher empirical likelihood than $\wreoptfull$.

Calculating the weights $\afull$ directly is infeasible because it requires access to the full dataset.
Fortunately, we don't need to consider all the data points for an accurate estimator.
The parameter space $\W$ is $m$-dimensional.
So intuitively, we only need $O(m)$ data points to solve such a problem to our desired accuracy.
%(This intuition is formalized in Section \ref{sec:anal}.)
This intuition motivates the following estimator.
Let $\Zreopt_i\subset Z_i$ be a set of $\nreopt$ data points uniformly sampled from $Z_i$ without replacement,
and let $\Zreopt$ be the union of the $\Zreopt_i$s.
Then the OWA estimator is defined as
\begin{equation}
\wreopt = \hat W \ahat
\end{equation}
where
\begin{equation}
\ahat = \argmax_\alpha \sum _{(\x,y)\in \Zreopt} f\left(y,\trans\x \hat W \alpha \right)
\end{equation}

Algorithm \ref{fig:alg2} shows the steps for efficiently computing $\wreopt$.
In the first round, each machine calculates $\wmle_i$ independently and broadcasts the results to every other machine.
Since the parameter vector has $d$ dimensions and there are $m$ machines, a total of $O(md)$ bits of get transmitted.
In the second round, each machine projects its local dataset $\Zreopt$ onto the space $\W$.
These projected data points are then transmitted to a predesignated master machine.
The projected data points each have dimension $m$, so a total of $O(m^2\nreopt)$ bits are transmitted.
The master machine now has all of the information to complete the optimization.

\begin{figure}
\hspace{-0.1in}
\begin{tikzpicture}
    [
    dot/.style = {minimum width=0.15cm,inner sep=0pt,line width=0pt,fill,circle,black,font=\small}
    ]
\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.5,-0.75) (-1.05,1) (-0.1,-0.1) (0.75,-0.5) (0.45,-0.7) };
\draw[gray] plot [smooth cycle,tension=1] coordinates {(-0.85,-0.85) (-1.3,1.55) (0.25,0) (2.5,-0.5) (1,-0.95)};
\draw[gray] plot [smooth cycle,tension=1] coordinates {(-1.15,-1.2) (-1.5,2) (1,0) (5,0) (3,-0.95)};

\draw[thick] (-2.2,2.15) -- (5.2,1.03);

\node[dot] (wstar) at (-0.28,-0.25) {};
\node at (0.05,-0.55) {$\wmle$};

@@

%\node[dot] (wavestar) at (4,0.5) {};
%\node                 at (4.5,0.3) {$\E\wave$};
\node[dot] (wave) at (3.8,1.25) {};
\node at (4.0,1.6) {$\wave$};
\node[dot] (wave) at (2.80,1.40) {};
\node at (3.1,1.75) {$\wmle_1$};
\node[dot] (wave) at (4.75,1.10) {};
\node at (4.9,1.45) {$\wmle_2$};
%\node[dot] (wave) at (3.8,1.25) {};

\node[dot] at (-1.35,2.03) {};
\node at (-0.9,2.4) {$\wreoptfull$};

\node at (4,-1.1) {$f(y,\trans\x\w)$};
\node at (5.65,1.0) {$\W$};

\end{tikzpicture}
\caption{
    Our method performs a second round of optimization to find the best parameter vector in $\W$.
    Since $\wave\in\W$, the estimator $\wreoptfull$ has higher empirical likelihood than $\wave$ almost surely.
}
\label{fig:contour}
\end{figure}

\begin{algorithm}[t]
\caption{Estimation in two rounds}
\label{alg:distributed}
\begin{algorithmic}
\State Each machine $i$ independently
\State ~~~~~loads dataset $Z_i$
\State ~~~~~calculates $\wmle_i$
\State All machines broadcast $\wmle_i$ to all other machines
\State Each machine $i$ independently
\State ~~~~~constructs $\hat W$
\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
\State ~~~~~calculates $\Zproj_i=\{(\trans\x\hat W,y) : (\x,y)\in\Zreopt\}$
\State All machines broadcast $\Zproj_i$ to a master machine
\State The master calculates $\wreopt$
\end{algorithmic}
\label{fig:alg2}
\end{algorithm}

%\begin{algorithm}[t]
%\caption{Estimation with common data}
%\label{alg:distributed}
%\begin{algorithmic}
%\State Each machine $i$ independently
%\State ~~~~~loads dataset $Z_i$
%\State ~~~~~calculates $\wmle_i$
%\State All machines broadcast $\wmle_i$ to all other machines
%\State Each machine $i$ independently
%\State ~~~~~constructs $\hat W$
%\State ~~~~~randomly selects a dataset $Z^{reopt}_i\subset Z_i$
%\State ~~~~~calculates $\trans\x\hat W$
%\State All machines broadcast $Z^{reopt}_i$ to a master machine
%\State The master calculates $\wreopt$
%\end{algorithmic}
%%\label{fig:alg}
%\end{algorithm}

@@

In this section, we analyze OWA's statistical performance.
The main tool we use is the Central Limit Theorem (CLT).
Informally, this theorem states that the $\wmle_i$ parameter vectors are approximately normally distributed.
%We then use results from the random projection literature to bound the distance
We then combine this result with standard results in random projections.
Classical versions of the CLT apply only in the limit as $n\to\infty$.
But we use a more advanced version of the CLT due to \cite{spokoiny2012parametricestimation} that applies in the finite sample case.
This approach makes our analysis simpler and more general than the analysis of algorithms discussed in Section \ref{sec:anal}.
All the technical details are handled by the CLT.
In particular, we prove novel bounds for the $\wave$ estimator.

\subsection {Subgaussian}

%We say that a random vector $X$ is \emph{subgaussian} with mean $\mu$ and covariance $\Sigma$
%if there exists a constant $c>0$ such that
%\begin{equation}
%\prob{\ltwo{\Sigma(X-\mu)} > t} \le c \exp(-t^2)
%\end{equation}
%The simplest example of a subgaussian random vector is a gaussian random vector.
%We denote this relationship in symbols as $X \lesssim \normal{\mu}{\Sigma}$.

We say that a random vector $\x\in\mathbb{R}^d$ is \emph{subgaussian} with variance proxy $\sigma>0$ if for all $\alpha\in\mathbb{R}^d$,
\begin{equation}
\E\exp(\trans\alpha\x) \le \exp(\ltwo{\alpha}^2\sigma^2/2)
.
\end{equation}
The simplest example of a subgaussian random variable is a multivariate normal vector with mean $\zero$ and covariance $I$.
In this case, $\sigma=1$.

\subsection{The Central Limit Theorem}

For completeness,
we state in this section a simplified version of Theorem 5.2 in \cite{spokoiny2012parametricestimation}.
We begin by stating the theorem's assumptions.
These assumptions are general.
We note in particular that logistic regression satisfies these assumptions,
and that these assumptions do not require the likelihood function be convex.

%\begin{enumerate}
%\item
\begin{assumption}
The likelihood function $f(y,\trans\x\w)$ is two times differentiable with respect to $\w$.
\end{assumption}
%\item
\begin{assumption}
For all $u>0$, there exists a function $g_1(u)>0$ such that for all $|\lambda| \le g_1(u)$
and for all $\x\in\X$,
\begin{equation}
\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} \le u}
\log\E\exp\left(\lambda\frac{\trans\x\nabla_\w f(y_i,\trans\x\w)}{\ltwo{V\x}}\right)
\le
\frac{v_0^2\lambda^2}{2}
\end{equation}
where $\trans V V = \I_\wstar$ and $v_0$ is a constant.
\end{assumption}
%\item
\begin{assumption}
For all $u>0$, there exists a function $b(u)>0$ such that
%\begin{equation}
%\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} = u}
%\log\E\exp\left(\frac{\E\left(f(y,\trans\x\w) - f(y,\trans\x\wstar)\right)}{\ltwo{\I_\wstar(\w-\wstar)}}\right)
%\le
%b(u)
%\end{equation}
\begin{equation}
\sup_{\w:\ltwo{\I_\wstar(\w-\wstar)} = u}
\log\E\exp\left(\frac{k(\w,\wstar)}{\ltwo{V(\w-\wstar)}}\right)
\le
b(u)
\end{equation}
where
%\begin{equation}
$
k(\w,\wstar) = \E\left(f(y,\trans\x\w) - f(y,\trans\x\wstar)\right)
$
.
%\end{equation}
\end{assumption}
%\end{enumerate}

\begin{theorem}
Assume Conditions 1-3 hold.
Then with probability at least $1-\delta$,
\begin{equation}
\ltwo{V(\w-\wstar)} \le \sqrt{\frac{d+\log{1/\delta}}{n}}
\end{equation}
\end{theorem}

@@

\subsection{In the Normal Regime}
%\label{sec:normal}
%
%\begin{theorem}
%We have that with probability at least
%\begin{equation}
%\wmle-\wstar \lesssim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
%\end{equation}
%In the limit as $n\to\infty$, we have the stronger statement that
%\begin{equation}
%\wmle-\wstar \sim \frac{1}{\sqrt{n}}\normal{\zero}{\I^{-1}_{\wstar}}
%\end{equation}
%\end{theorem}

@@

We assume that for every estimator $\hat\w$ trained on $n$ data points,
\begin{equation}
\label{eq:assumption}
\hat\w - \E\hat\w \sim \frac{1}{\sqrt n}\normal{\zero}{\I^{-1}_{\E\hat\w}}
%\hat\w \sim \normal{\E\hat\w}{\frac{1}{n}I^{-1}_{\E\hat\w}}
\end{equation}
%
%Throughout Section \ref{sec:normal} we assume that $\wave_i$ is normally distributed with mean $\wtave$ and covariance $\frac{1}{n}I^{-1}_{\wtave}$.
%\begin{equation}
%\label{eq:assumption}
%\wmle_i \sim \normal{\E\wmle_i}{\frac{1}{n}I^{-1}_{\E\wmle_i}}
%\end{equation}
%Here, $I_{\wtave}$ is the Fisher information matrix at point $\wtave$.
Here, $\I_{\E\hat\w}$ is the Fisher information matrix for parameter $\E\hat\w$.
This assumption is motivated by the central limit theorem for maximum likelihood estimation.
%There are many versions of this theorem.
Theorem 7.5.2 of \cite{lehmann1999elements} is particularly simple version and reproduced below.
\begin{theorem}
\label{thm:clt}
Assume the following mild regularity conditions:
\begin{enumerate}[noitemsep,topsep=0pt]%,label=(\itshape\roman*)]
\item all data points in $Z_i$ are i.i.d.;
\item the parameter space $\Theta$ is open;
\item the model is identifiable and consistent;
\item the matrix $\I_{\wtave}$ is non-singular;
\item the support set $A = \{(y,\x) : f(y,\trans\x\theta) > 0\}$ is independent of $\theta$;
and \item for all $(y,\x)\in A$, $f(y,\trans\x\theta)$ is three times differentiable with respect to $\theta$,
and the third derivative is continuous.
\end{enumerate}
Then in the limit as $n\to\infty$,
\begin{equation*}
\wave_i \law \wtave + \frac{1}{\sqrt n} \normal{0}{\I^{-1}_\wstar}
\end{equation*}
Here the notation $\law$ denotes convergence in law,
which is the standard notion of convergence for a sequence of probability distributions.
\end{theorem}

@@

There are many generalizations of Theorem \ref{thm:clt}.
For example, Theorem 4.2 in \cite{spokoiny2012parametricestimation} provides finite sample bounds with explicit dependence on the degree of model misspecification, optimization error, and the lack of independence in the data variables.
Because Spokoiny's theorem is non-asymptotic,
the distribution of the estimator is a sub-Gaussian random variable.
This more general setting requires more extensive notation,
but all of our results hold.

@@

Note that this reduction in variance is essentially optimal.
By an argument similar to Equations \ref{eq:var1}-\ref{eq:var2},
we can show that the variance of the maximum likelihood estimator trained on all of the data is
\begin{equation}
\wmle - \E\wmle \sim \frac{1}{\sqrt{mn}} \normal{\zero}{\I^{-1}_{\wmle}}
\label{eq:var3}
\end{equation}
The only difference between Equations \ref{eq:var2} and \ref{eq:var3} is the location of the Fisher information matrix.

@@

\begin{proof}
This proof follows the same form as the proof for Theorem \ref{thm:wave}.
We first decompose the error into a bias and variance component;
then we bound each component separately.
To decompose the error, we need to introduce the following affine space.
\begin{align}
\Waff&=\affspan \{\wmle_i\}_{i=1}^m
\\
\WaffE&=\affspan \{\wmle_i - \wave + \E\wave\}_{i=1}^m
\end{align}
We now have that
\begin{align}
\ltwo{\wstar-\proj\W\wstar}
&\le
\ltwo{\wstar-\proj\Waff\wstar}
\\
&\le
\ltwo{\wstar-\proj\WaffE\wstar}
%\\&
+
\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
%\\&+
%\ltwo{\proj\Wtave\wstar-\proj\Wave\wstar}
\end{align}
The first line above follows because $\Waff\subseteq\W$,
and the second by the triangle inequality.
The left term of the inequality above can be thought of as akin to the bias of the estimator $\proj\Waff\wstar$,
and the right term can be thought of as akin to the variance.

We begin by bounding the right term.
We have that with probability at least $1-\exp(-t)$,
\begin{align}
&\ltwo{\proj\WaffE\wstar-\proj\Waff\wstar}
\\
&=
\ltwo{(\proj\Waff\wstar+\wave-\wtave)-\proj\Waff\wstar}
\\&=
\ltwo{\wave-\wtave}
\\&\le
\sqrt\frac{v_t}{mn}
\end{align}
The first equality follows because the space $\WaffE$ is a translation of $\Waff$ by the vector $(\wave-\wtave)$.
The probabilistic inequality is due to Equation \ref{eq:ct1} in the proof of Theorem \ref{thm:wave}.
